The Revisited Hidden Weight Bit Function

### 2025-03-26 RevisedHiddenWeight.tex
Notation modif's

\def\anon{1} %Set to 1 for anonymous submissions, hides acknowledgements and author names
\def\full{1} %Set to 0 for springer proceedings

@@

These various results on the algebraic properties of the revisited \hwbf{} suggest that it possesses strong resistance against standard attacks. Furthermore, its high nonlinearity makes it a well-suited candidate for use as a filter function in contexts such as filtered \textsf{LFSR}s or in homomorphically-friendly schemes like \textsf{FLIP} and \textsf{FiLIP}.

\pmnote{For example, the filter functions leading to the best current performances for FiLIP in HHE~\cite{CCS:CDPP22,CIC:MeaParPei24,eprint:AGHM24} are xor-threshold functions, the direct sum of a $k$-variable linear function and an $m$-variable function giving $1$ if and only if the input has Hamming weight at least a threshold $d$ (which generalizes the majority function). Using Proposition $7$ in~\cite{IEEE:CarMea21}, the nonlinearity of a xor-threshold function is at most:
\[2^{n-1}-2^k \cdot M \text{ where }
M=\left\{\begin{array}{ll}
\binom{m-1}{\frac{m-1}{2}}&\text{ if }m\equiv 1\bmod 2,\\
\frac{1}{2} \binom{m}{\frac{m}{2}}&\text{ if }m\equiv 0\bmod 2.\\
\end{array}\right.
\]
The nonlinearity of this whole family of functions is similar to the one of the majority, hence the revisited HWBF provides a better nonlinearity that currently used filters for FiLIP for the same sizes, allowing to thwart correlation-like attacks with functions in less variables.}

\begin{table}
	\scriptsize
	\centering
	\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
		\hline
		$ n $      & $4$     & $6$     & $8$     & $10$    & $12$    & $14$    & $16$    & $18$    \\
		\hline
		$(d, e)$ & $(1,2)$ & $(1,3)$ & $(1,5)$  & $(1,7)$ & $(1,9)$ & $(1,11)$ & $(1,13)$ & $(1,15)$ \\ 
		&     &    & $(2,4)$  & $(2,6)$ & $(2,8)$ & $(2,10)$ & $(2,12)$ & $(2,14)$\\ 
		&     &    &     & $(3,5)$ & $(3,6)$ & $(3,9)$ & $(3,11)$ & $(3,13)$\\ 
		&     &    &     &  &  &  & $(4,9)$ & $(4,11)$\\ 
	 	&     &    &     &  &  &  &  & $(5,10)$ \\
		\hline
	\end{tabular}
	\mycaption{Lowest possible values of $(d, e)$ for different values of $n$.}
	\label{table:FAI}
\end{table}

>>>>

\def\anon{0} %Set to 1 for anonymous submissions, hides acknowledgements and author names
\def\full{1} %Set to 0 for springer proceedings

@@

These various results on the algebraic properties of the revisited \hwbf{} suggest that it possesses strong resistance against standard attacks. Furthermore, its high nonlinearity makes it a well-suited candidate for use as a filter function in contexts such as filtered \textsf{LFSR}s or in homomorphically-friendly schemes like \textsf{FLIP} and \textsf{FiLIP}. For example, the best current \textsf{FiLIP} filters in \textsf{HHE}~\cite{CCS:CDPP22,CIC:MeaParPei24,eprint:AGHM24} are \textsf{XOR}-threshold functions—the sum of a $k$-variable linear function and an $m$-variable threshold function (generalizing majority).  
By Proposition $7$ in~\cite{IEEE:CarMea21}, their nonlinearity is at most:
\[
2^{k+m-1} - 2^k \cdot M, \quad \text{where } M = \begin{cases}
\binom{m-1}{\frac{m-1}{2}} & \text{if } m \text{ is odd}, \\
\frac{1}{2} \binom{m}{\frac{m}{2}} & \text{if } m \text{ is even}.
\end{cases}
\]
This is similar to the majority function, whereas the revisited \hwbf{} achieves higher nonlinearity for the same size, offering better resistance to correlation-like attacks with fewer variables.
\begin{table}
	\scriptsize
	\centering
	\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
		\hline
		$ n $      & $4$     & $6$     & $8$     & $10$    & $12$    & $14$    & $16$    & $18$    \\
		\hline
		$(d, e)$ & $(1,2)$ & $(1,3)$ & $(1,5)$  & $(1,7)$ & $(1,9)$ & $(1,11)$ & $(1,13)$ & $(1,15)$ \\ 
		&     &    & $(2,4)$  & $(2,6)$ & $(2,8)$ & $(2,10)$ & $(2,12)$ & $(2,14)$\\ 
		&     &    &     & $(3,5)$ & $(3,6)$ & $(3,9)$ & $(3,11)$ & $(3,13)$\\ 
		&     &    &     &  &  &  & $(4,9)$ & $(4,11)$\\ 
	 	&     &    &     &  &  &  &  & $(5,10)$ \\
		\hline
	\end{tabular}
	\mycaption{Lowest possible values of $(d, e)$ for different values of $n$.}
	\label{table:FAI}
\end{table}
